---
title: "Comparison_Script"
author: "Anusuiya Bora; s222331482@deakin.edu.au"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## For PMID36259947 using DAVID2023q4

### Loading libraries:

```{r, libraries}

library("knitr")
library("kableExtra")
library("tidyverse")
library("stringr")
library("tidyr")
library("eulerr")

```

### DAVID Results from the web-tool (2023q4):

```{r, setupDAVIDresults}

DAVIDweb_results <- read.table("../output_lists/PMID36259947_results.txt", header = TRUE, sep = "\t")  #Change PMID/PMCID

#PMID36259947_results <- read.csv("../output_lists/PMID33947848_webDAVID_20240513.tsv", header = TRUE, sep = "\t")

split_terms <- str_split(DAVIDweb_results$Term, pattern = "~")

DAVIDweb_results$GOID <- sapply(split_terms, "[", 1)
DAVIDweb_results$GO.Term <- sapply(split_terms, "[", 2)

DAVIDweb_results <- select(DAVIDweb_results, GOID, GO.Term, Category, Count, X., PValue, Genes, List.Total, Pop.Hits, Pop.Total, Fold.Enrichment, Bonferroni, Benjamini, FDR)

tmp <- DAVIDweb_results
tmp$Genes=NULL

head(tmp,50) %>%      #Change PMID/PMCID
  kbl(caption="Top results from the DAVID Website") %>%
  kable_paper("hover", full_width = F)

```

### DAVID Results from our R script/DAVID Reborn Tool:

```{r DAVID_Reborn results}

script_results <- readRDS("../DAVID_vScripts/ora_arranged2023.RDS")

script_results$GOID <- sapply(script_results$Description, function(term) strsplit(term, " ")[[1]][1])

script_results$GO.Term <- sapply(script_results$Description, function(term) paste(strsplit(term, " ")[[1]][-1], collapse = " "))

script_results <- select(script_results, S_No, GOID, GO.Term, GeneRatio, BgRatio, pvalue, p.adjust, qvalue, geneID, Count, gr, br, EnrichmentScore)

# minimum of 2 genes
#script_results <- subset(script_results,Count >=2 )

tmp <- script_results
tmp$geneID=NULL

head(tmp,50) %>%
  kbl(caption="Top results from DAVID Reborn Tool") %>%
  kable_paper("hover", full_width = F)

```

## Compare the genes that are recognised in both analyses

Extract the gene lists from the results of the web tool and the reborn tool to see where genes are being exluded.

```{r,geneoverlap1}

web_genes <- unique(gsub(" ","",unlist(strsplit(DAVIDweb_results$Genes,","))))
web_genes <- sort(web_genes)
str(web_genes)
#481 genes

script_genes <- unique(unlist(strsplit(script_results$geneID," ")))
script_genes <- sort(script_genes)
str(script_genes)
# 499 genes looks right

v1 <- list("web"=web_genes,"script"=script_genes)

plot(euler(v1),quantities = list(cex = 1.0), labels = list(cex = 1.5))

message("Genes unique to web")
setdiff(web_genes,script_genes)

message("Genes unique to script")
setdiff(script_genes, web_genes)

```

## Comparing results from the web-server and our R script:

```{r,comparison1}

# From DAVID webserver table output: 
#web_split <- sapply(strsplit(PMID33947848_results$Term,"~"),"[[",1)

DAVID_webserver <- DAVIDweb_results %>% #Change PMID/PMCID
  select(GOID, Count, PValue, Fold.Enrichment, FDR)

DAVID_reborn <- script_results %>%
  select(S_No, GOID, GO.Term, GeneRatio, pvalue, p.adjust, Count ,EnrichmentScore)

comparison_table <- merge(DAVID_webserver, DAVID_reborn, by = "GOID", all = TRUE)

comparison_table <- comparison_table[order(comparison_table$PValue),]

comparison_table$geneID=NULL

head(comparison_table,50) %>%
  kbl(caption="Top replication results") %>%
  kable_paper("hover", full_width = F)

```

Compare p-values directly.

```{r,comparison2}

comparison_table$pvaluediff <- abs((comparison_table$PValue) - (comparison_table$pvalue))

hist( comparison_table$pvaluediff / (comparison_table$PValue) )

table( comparison_table$pvaluediff / (comparison_table$PValue) < 0.5)

comparison_table$pvaluecheck  <-  comparison_table$pvaluediff / (comparison_table$PValue) < 0.5

```

Compare adjusted p-values directly.

```{r,comparison3}

comparison_table$fdrdiff <- abs(-log10(comparison_table$FDR) - -log10(comparison_table$p.adjust))

hist( comparison_table$fdrdiff / -log10(comparison_table$FDR) )

table(comparison_table$fdrdiff / -log10(comparison_table$FDR) < 1)

comparison_table$fdrcheck  <- comparison_table$fdrdiff / -log10(comparison_table$FDR) < 1

```

Compare fold enrichments directly.

```{r,comparison4}

comparison_table$folddiff <- abs( comparison_table$Fold.Enrichment - comparison_table$EnrichmentScore)

hist( comparison_table$folddiff / comparison_table$Fold.Enrichment )

table( comparison_table$folddiff / comparison_table$Fold.Enrichment <0.3)

comparison_table$foldcheck  <- comparison_table$folddiff / comparison_table$Fold.Enrichment <0.3

```

Now look at the significant sets and see if they classify as replicated.

```{r,comparison5}

sig <- subset(comparison_table, FDR <0.05)

repro <- subset(sig, pvaluecheck == "TRUE")

#repro <- subset(sig, fdrcheck == "TRUE" & foldcheck == "TRUE")

notrepro <- subset(sig, pvaluecheck != "TRUE")

message("No. significant results from paper")
nrow(sig)

message("No. replicated results")
nrow(repro)

message("No. non-replicated results")
nrow(notrepro)

message("Success rate")
nrow(repro) / nrow(sig)

```

```{r,tables}

head(repro,50) %>%
  kbl(caption="Top reproduced results") %>%
  kable_paper("hover", full_width = F)

head(notrepro,50) %>%
  kbl(caption="Top un-reproduced results") %>%
  kable_paper("hover", full_width = F)

```

## Session information

For reproducibility

```{r,session}

sessionInfo()

```
