---
title: "DAVID Redone Code"
author: "Anusuiya Bora; s222331482@deakin.edu.au"
#output: html_document
output: 
  html_document:
    code_folding: show 
date: "2024-05-01"
---

### This is for DAVID 2021 (May-June 2022, before it got updated to latest versions)

### Methodology:
1. Loading libraries

```{r loading_libraries}

library("clusterProfiler")
library("kableExtra")
library("dplyr")
library("rio")
library("here")
library("gtsummary")
library("ggplot2")

```


2) Input gene list:

``` {r loading_input_genes}

#genelist_path <- "~/Projects/DAVID_Reborn/Data/input_genelist"

genelist <- read.csv("../input_lists/PMC6587650.csv")

Genevec <- genelist[, 1]

```

3) Loading GO datsets

```{r include=FALSE, echo=FALSE, load_GORDS}

gobp2021 <- readRDS("../go_gaf_datasets/go2021/GO_bp_2021.Rds")
gocc2021 <- readRDS("../go_gaf_datasets/go2021/GO_cc_2021.Rds")
gomf2021 <- readRDS("../go_gaf_datasets/go2021/GO_mf_2021.Rds")
gocombined2021 <- readRDS("../go_gaf_datasets/go2021/GO_combined_2021.Rds")

```


```{r include=FALSE, echo=FALSE, converting_RDS2gmt}
gobp_terms <- rev(stack(gobp2021))
str(gobp_terms)

gocc_terms <- rev(stack(gocc2021))
str(gocc_terms)

gomf_terms <- rev(stack(gomf2021))
str(gomf_terms)

gocombined_terms <- rev(stack(gocombined2021))
str(gocombined_terms)

#goterms <- rev(stack(golist))
#str(goterms)

```

5) Performing Enrichment analysis

```{r include=T, echo=T, enricher_code}
ora <- as.data.frame(enricher(gene = Genevec ,
                              minGSSize = 10,
                              maxGSSize = 50000, 
                              TERM2GENE = gocombined_terms, #for TERM2GENE, options are: gobp_terms, gocc_terms, gomf_terms, gocombined_terms
                              pAdjustMethod="fdr",  
                              pvalueCutoff = 1, 
                              qvalueCutoff = 1  ))

x <- gsub("/", " ", ora$geneID) #to separate geneIds from /

head(x)

dim(ora)

ora$geneID <- x

ora$ID <- NULL
rownames(ora) <- NULL #removing empty 1st column name

ora$S_No <- seq_len(nrow(ora))
ora <- ora[, c("S_No", names(ora)[-which(names(ora) == "S_No")])]

head(ora)

```


6) Calculating enrichment scores

``` {r include=FALSE, echo=FALSE, ES_calculation}

ora$gene <- NULL

gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

es <- gr/br

ora_new <- cbind(ora, gr, br, es)

names(ora_new)[names(ora_new) == "es"] <- "EnrichmentScore"

ora_arranged2021 <- ora_new %>% 
  arrange(p.adjust)

head(ora_arranged2021)


```

7) Selecting pathways with FDR<0.05

```{r pathways_FDR}

ora_fdr2021 <- ora_arranged2021 %>%
  filter(p.adjust < 0.05)

```



```{r output_table}

ora_fdr2021 %>%
  kbl(caption="Genes and GO Terms identified, FDR &lt; 0.05") %>%
  kable_paper("hover", full_width = F)

```

```{r complete_results, results='asis', echo=FALSE}

HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )
 
cat(HEADER)
 
kbl(ora_arranged2021, format="html", row.names=FALSE, caption = "Complete results",digits=100) %>%
kable_styling("hover", full_width = FALSE)
 
cat("<br></p></details>")
cat("<hr><br>")

```


### Save object

```{r save_object}

saveRDS(ora_arranged2021, file = "ora_arranged2021.RDS")
saveRDS(ora_fdr2021, file = "ora_fdr2021.RDS")
# z <- readRDS("ora.RDS")  #To read RDS 

```

### Session information

For reproducibility.

```{r,sessioninfo}

sessionInfo()

```
