---
title: "DAVID Redone Code"
author: "Anusuiya Bora; s222331482@deakin.edu.au"
#output: html_document
output: 
  html_document:
    code_folding: show 
date: "2024-05-01"
---

### This is for DAVID 2023 - quarter 4 version

### Methodology:
1. Loading libraries

```{r loading_libraries}

library("clusterProfiler")
library("kableExtra")
library("dplyr")
library("rio")
library("here")
library("gtsummary")
library("ggplot2")

```


2) Input gene list:

``` {r loading_input_genes}

#genelist_path <- "~/Projects/DAVID_Reborn/Data/input_genelist"

genelist <- read.csv("../input_lists/PMC6587650.csv")

# getwd()
# setwd ("..")
#genelist <- "PMC6587650.csv"
#genelist <- read.csv("PMC6587650.csv")

Genevec <- genelist[, 1]

```

3) Loading GO datsets

```{r include=FALSE, echo=FALSE, load_GORDS}


gobp_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_bp_2023q4.Rds")

# getwd()
# setwd(..) 

gomf_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_mf_2023q4.Rds")

gocc_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_cc_2023q4.Rds")

gocombined_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_combined_2023q4.Rds")

# gobp <- rio::import(here::here("Data", "GO_dataset", "GObp.rds"))
# gocc <- rio::import(here::here("Data", "GO_dataset", "GOcc.rds"))
# gomf <- rio::import(here::here("Data", "GO_dataset", "GOmf.rds"))
# gocombined <- rio::import(here::here("Data", "GO_dataset", "GOcombined.rds"))

```


```{r include=FALSE, echo=FALSE, converting_RDS2gmt}
gobp_terms <- rev(stack(gobp_2023))
str(gobp_terms)

gocc_terms <- rev(stack(gocc_2023))
str(gocc_terms)

gomf_terms <- rev(stack(gomf_2023))
str(gomf_terms)

gocombined_terms <- rev(stack(gocombined_2023))
str(gocombined_terms)

#goterms <- rev(stack(golist))
#str(goterms)

```

5) Performing Enrichment analysis

```{r include=T, echo=T, enricher_code}
ora <- as.data.frame(enricher(gene = Genevec ,
                              minGSSize = 10,
                              maxGSSize = 50000, 
                              TERM2GENE = gocombined_terms, #for TERM2GENE, options are: gobp_terms, gocc_terms, gomf_terms, gocombined_terms
                              pAdjustMethod="fdr",  
                              pvalueCutoff = 1, 
                              qvalueCutoff = 1  ))

x <- gsub("/", " ", ora$geneID) #to separate geneIds from /

head(x)

dim(ora)

ora$geneID <- x

ora$ID <- NULL
rownames(ora) <- NULL #removing empty 1st column name

ora$S_No <- seq_len(nrow(ora))
ora <- ora[, c("S_No", names(ora)[-which(names(ora) == "S_No")])]

head(ora)

```


6) Calculating enrichment scores

``` {r include=FALSE, echo=FALSE, ES_calculation}

ora$gene <- NULL

gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

es <- gr/br

ora_new <- cbind(ora, gr, br, es)

names(ora_new)[names(ora_new) == "es"] <- "EnrichmentScore"

ora_arranged <- ora_new %>% 
  arrange(p.adjust)

head(ora_arranged)


```

7) Selecting pathways with FDR<0.05

```{r pathways_FDR}

ora_fdr2023 <- ora_arranged %>%
  filter(p.adjust < 0.05)

```



```{r output_table}

ora_fdr2023 %>%
  kbl(caption="Genes and GO Terms identified, FDR &lt; 0.05") %>%
  kable_paper("hover", full_width = F)

```

```{r complete_results, results='asis', echo=FALSE}

HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )
 
cat(HEADER)
 
kbl(ora_arranged, format="html", row.names=FALSE, caption = "Complete results",digits=100) %>%
kable_styling("hover", full_width = FALSE)
 
cat("<br></p></details>")
cat("<hr><br>")

```

```{r pie_chart}

# aggregate_data <- aggregate(EnrichmentScore ~ Description, data = ora_fdr, FUN = sum)

# Create a pie chart
# pie(aggregate_data$EnrichmentScore, labels = aggregate_data$Description)

# ggplot(ora_fdr, aes(x = "", y = Description, fill = EnrichmentScore)) +
#  geom_bar(stat = "identity") +
#  coord_polar(theta = "y") +
#  labs(title = "Pie Chart for enriched terms", fill = "EnrichmentScore", y = NULL) +
#  theme_minimal()

```

### Save object

```{r save_object}

saveRDS(ora_arranged, file = "ora_arranged2023.RDS")
saveRDS(ora_fdr2023, file = "ora_fdr2023.RDS")
# z <- readRDS("ora.RDS")  #To read RDS 

```

### Session information

For reproducibility.

```{r,sessioninfo}

sessionInfo()

```
