---
title: "DAVID Redone Code"
author: "Anusuiya Bora; s222331482@deakin.edu.au"
output:
  html_document:
    code_folding: show
date: "2024-05-01"
---

This is for DAVID 2023 - quarter 4 version

### Loading libraries

```{r,loading_libraries}

library("clusterProfiler")
library("kableExtra")
library("dplyr")
library("rio")
library("here")
library("gtsummary")
library("ggplot2")

```


### Input gene list

```{r,loading_input_genes}

Genevec <- read.table("../input_lists/PMID33947848_genes.txt")[,1]

```

### Loading GO datsets

```{r,load_GORDS}

gobp_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_bp_2023q4.Rds")

gomf_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_mf_2023q4.Rds")

gocc_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_cc_2023q4.Rds")

gocombined_2023 <- readRDS("../go_gaf_datasets/go2023q4/GO_combined_2023q4.Rds")

```

### Convert gene lists to dataframe for clusterprofiler

```{r,converting_RDS2gmt}

gobp_terms <- rev(stack(gobp_2023))
str(gobp_terms)

gocc_terms <- rev(stack(gocc_2023))
str(gocc_terms)

gomf_terms <- rev(stack(gomf_2023))
str(gomf_terms)

gocombined_terms <- rev(stack(gocombined_2023))
str(gocombined_terms)

```

### Performing Enrichment analysis

for TERM2GENE parameter of clusterprofiler the options are: gobp_terms, gocc_terms, gomf_terms, gocombined_terms

```{r,enricher_code}

ora <- as.data.frame(enricher(gene = Genevec ,
                              minGSSize = 1,
                              maxGSSize = 50000,
                              TERM2GENE = gocombined_terms,
                              pAdjustMethod="fdr",
                              pvalueCutoff = 1,
                              qvalueCutoff = 1  ))

ora$geneID <- gsub("/", " ", ora$geneID) #to separate geneIds from /

ora$ID <- NULL
rownames(ora) <- NULL #removing empty 1st column name

ora$S_No <- seq_len(nrow(ora))
ora <- ora[, c("S_No", names(ora)[-which(names(ora) == "S_No")])]

head(ora)

```

Now test the presence of a few gene sets.

GO:0001654

GO:0001764

GO:0002682

```{r,test1}

# collect reference sets
GO0001654 <- gocombined_terms[grep("GO:0001654",gocombined_terms$ind),2]
GO0001764 <- gocombined_terms[grep("GO:0001764",gocombined_terms$ind),2]
GO0002682 <- gocombined_terms[grep("GO:0002682",gocombined_terms$ind),2]

# intersect with input list
length(which(GO0001654 %in% Genevec))
length(which(GO0001764 %in% Genevec))
length(which(GO0002682 %in% Genevec))

```

### Calculating enrichment scores

```{r,ES_calculation}

ora$gene <- NULL

gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

es <- gr/br

ora_new <- cbind(ora, gr, br, es)

names(ora_new)[names(ora_new) == "es"] <- "EnrichmentScore"

ora_arranged2023 <- ora_new %>%  #ora_arranged naming done after pathways were arranged from ascending order of FDR 
  arrange(p.adjust)

head(ora_arranged2023)

```

### Top table

```{r,output_table}

head(ora_arranged2023,50) %>%
  kbl(caption="Top 50 GO Terms identified") %>%
  kable_paper("hover", full_width = F)

```

```{r,complete_results, results='asis', echo=FALSE}

HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )

cat(HEADER)

kbl(ora_arranged2023, format="html", row.names=FALSE, caption = "Complete results",digits=100) %>%
kable_styling("hover", full_width = FALSE)

cat("<br></p></details>")
cat("<hr><br>")

```

### Save object

```{r,save_object}

saveRDS(ora_arranged2023, file = "ora_arranged2023.RDS")
# z <- readRDS("ora.RDS")  #To read RDS 

```

### Session information

For reproducibility.

```{r,sessioninfo}

sessionInfo()

```
