---
title: "DAVID Redone Code"
author: "Anusuiya Bora; s222331482@deakin.edu.au"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
date: "`r Sys.Date()`"
---

## This is for DAVID 2016 (6.8 version) 

### Loading libraries

```{r, loading_libraries}

library("clusterProfiler")
library("kableExtra")
library("dplyr")
library("rio")
library("here")
library("gtsummary")
library("ggplot2")

```


### Input gene list:

``` {r, loading_input_genes}

genelist <- read.table("../input_lists/PMID36259947_genes.txt")

#getwd() 
#setwd("..")

Genevec <- genelist[, 1]

```

### Loading GO datsets

```{r, include=FALSE, echo=FALSE, load_GORDS}

gobp2016 <- readRDS("../go_gaf_datasets/go2016/GObp.rds")
gocc2016 <- readRDS("../go_gaf_datasets/go2016/GOcc.rds")
gomf2016 <- readRDS("../go_gaf_datasets/go2016/GOmf.rds")
gocombined2016 <- readRDS("../go_gaf_datasets/go2016/GOcombined.rds")

```

### Convert gene lists to dataframe for clusterprofiler

```{r, include=FALSE, echo=FALSE, converting_RDS2gmt}
gobp_terms <- rev(stack(gobp2016))
str(gobp_terms)

gocc_terms <- rev(stack(gocc2016))
str(gocc_terms)

gomf_terms <- rev(stack(gomf2016))
str(gomf_terms)

gocombined_terms <- rev(stack(gocombined2016))
str(gocombined_terms)

```

### Performing Enrichment analysis

for TERM2GENE, options are: gobp_terms, gocc_terms, gomf_terms, gocombined_terms

```{r, include=T, echo=T, enricher_code}
ora <- as.data.frame(enricher(gene = Genevec ,
                              minGSSize = 10,
                              maxGSSize = 50000, 
                              TERM2GENE = gocombined_terms, 
                              pAdjustMethod="fdr",  
                              pvalueCutoff = 1, 
                              qvalueCutoff = 1  ))

ora$geneID <- gsub("/", " ", ora$geneID) #to separate geneIds from /

dim(ora)


ora$ID <- NULL
rownames(ora) <- NULL #removing empty 1st column name

ora$S_No <- seq_len(nrow(ora))
ora <- ora[, c("S_No", names(ora)[-which(names(ora) == "S_No")])]

head(ora)

```

Now test the presence of a few gene sets.

GO:0001654

GO:0001764

GO:0002682

```{r,test1}

# collect reference sets
GO0001654 <- gocombined_terms[grep("GO:0001654",gocombined_terms$ind),2]
GO0001764 <- gocombined_terms[grep("GO:0001764",gocombined_terms$ind),2]
GO0002682 <- gocombined_terms[grep("GO:0002682",gocombined_terms$ind),2]

# intersect with input list
length(which(GO0001654 %in% Genevec))
length(which(GO0001764 %in% Genevec))
length(which(GO0002682 %in% Genevec))

```

### Calculating enrichment scores

``` {r, include=FALSE, echo=FALSE, ES_calculation}

ora$gene <- NULL

gr <- as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$GeneRatio,"/"),"[[",2))

br <- as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",1)) /
  as.numeric(sapply(strsplit(ora$BgRatio,"/"),"[[",2))

es <- gr/br

ora_new <- cbind(ora, gr, br, es)

names(ora_new)[names(ora_new) == "es"] <- "EnrichmentScore"

ora_arranged2016 <- ora_new %>%  #ora_arranged naming done after pathways were arranged from ascending order of FDR 
  arrange(p.adjust)

head(ora_arranged2016)

#ora_arranged2016 <- ora_new %>% 
#  arrange(p.adjust)

#head(ora_arranged2016)


```

### Selecting pathways with FDR<0.05

```{r, pathways_FDR}

#ora_fdr2016 <- ora_arranged2016 %>%
#  filter(p.adjust < 0.05)

```

### Top table

```{r,output_table}

head(ora_arranged2016,50) %>%
  kbl(caption="Top 50 GO Terms identified") %>%
  kable_paper("hover", full_width = F)

```

```{r,complete_results, results='asis', echo=FALSE}

HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )

cat(HEADER)

kbl(ora_arranged2016, format="html", row.names=FALSE, caption = "Complete results",digits=100) %>%
kable_styling("hover", full_width = FALSE)

cat("<br></p></details>")
cat("<hr><br>")

```


### Save object

```{r, save_object}

#saveRDS(ora_arranged2016, file = "ora.RDS")
saveRDS(ora_arranged2016, file = "ora_arranged2016.RDS")
#saveRDS(ora_fdr2016, file = "ora_fdr2016.RDS")
# z <- readRDS("ora.RDS")  #To read RDS 

```

### Session information

For reproducibility.

```{r,sessioninfo}

sessionInfo()

```
